---
title: "Supervised learning in R: Regression"
Author: Leyli Garryyeva
Platform: DataCamp
Course: CS688 - Data Mining
Assignment: 1
output: html_notebook
---
1. Linear Regression

Pros: 
  - Easy to fit and to apply
  - Concise
  - Less prone to overfitting (performance on new data is usually quite     similar to the performance on the training data)
  - Interpretable

Cons:
  - Can only exress linear and additive relationships
  - Collinearity -- when input variables are partially correlated
  - Coefficients might change sign
  - High collinearity:
    -- Coefficients (or standard errors) look too large
    -- Model may be unstable
  
----------------------------------------
Code a simple one-variable regression

For the first coding exercise, you'll create a formula to define a one-variable modeling task, and then fit a linear model to the data. You are given the rates of male and female unemployment in the United States over several years (Source).

The task is to predict the rate of female unemployment from the observed rate of male unemployment. The outcome is female_unemployment, and the input is male_unemployment.

The sign of the variable coefficient tells you whether the outcome increases (+) or decreases (-) as the variable increases.

Recall the calling interface for lm() is:
lm(formula, data = ___)

```{r}
# unemployment is loaded in the workspace
summary(unemployment)

# Define a formula to express female_unemployment as a function of male_unemployment
fmla <- female_unemployment ~ male_unemployment

# Print it
fmla

# Use the formula to fit a model: unemployment_model
unemployment_model <- lm(fmla, data = unemployment)

# Print it
unemployment_model
```

Examining a model

Let's look at the model unemployment_model that you have just created. There are a variety of different ways to examine a model; each way provides different information. We will use summary(), broom::glance(), and sigr::wrapFTest().

```{r}
# broom and sigr are already loaded in your workspace
# Print unemployment_model
unemployment_model

# Call summary() on unemployment_model to get more details
summary(unemployment_model)

# Call glance() on unemployment_model to see the details in a tidier form
glance(unemployment_model)

# Call wrapFTest() on unemployment_model to see the most relevant details
wrapFTest(unemployment_model)

```

Predicting once you fit the model:

Predicting from the unemployment model

In this exercise, you will use your unemployment model unemployment_model to make predictions from the unemployment data, and compare predicted female unemployment rates to the actual observed female unemployment rates on the training data, unemployment. You will also use your model to predict on the new data in newrates, which consists of only one observation, where male unemployment is 5%.

The predict() interface for lm models takes the form
          predict(model, newdata)

You will use the ggplot2 package to make the plots, so you will add the prediction column to the unemployment data frame. You will plot outcome versus prediction, and compare them to the line that represents perfect predictions (that is when the outcome is equal to the predicted value).

The ggplot2 command to plot a scatterplot of dframe$outcome versus dframe$pred (pred on the x axis, outcome on the y axis), along with a blue line where outcome == pred is as follows:

ggplot(dframe, aes(x = pred, y = outcome)) + 
       geom_point() +  
       geom_abline(color = "blue")


```{r}
# unemployment is in your workspace
summary(unemployment)

# newrates is in your workspace
newrates

# Predict female unemployment in the unemployment data set
unemployment$prediction <-  predict(unemployment_model)

# load the ggplot2 package
library(ggplot2)

# Make a plot to compare predictions to actual (prediction on x axis). 
ggplot(unemployment, aes(x = prediction, y = female_unemployment)) + 
  geom_point() +
  geom_abline(color = "blue")

# Predict female unemployment rate when male unemployment is 5%
pred <- predict(unemployment_model, newrates)
# Print it
pred
```

Multivariate linear regression (Part 1)

In this exercise, you will work with the blood pressure dataset (Source), and model blood_pressure as a function of weight and age.

Source: http://college.cengage.com/mathematics/brase/understandable_statistics/7e/students/datasets/mlr/frames/frame.html

Exercise: The data frame bloodpressure is in the workspace.

Define a formula that expresses blood_pressure explicitly as a function of age and weight. Assign the formula to the variable fmla and print it.
Use fmla to fit a linear model to predict blood_pressure from age and weight in the data set bloodpressure. Call the model bloodpressure_model.
Print the model and call summary() on it. Does blood pressure increase or decrease with age? With weight?


```{r}
# bloodpressure is in the workspace
summary(bloodpressure)

# Create the formula and print it
fmla <- blood_pressure ~ age + weight 
fmla

# Fit the model: bloodpressure_model
bloodpressure_model <- lm(fmla, bloodpressure)

# Print bloodpressure_model and call summary() 
bloodpressure_model
summary(bloodpressure_model)
```

Multivariate linear regression (Part 2)

Now you will make predictions using the blood pressure model bloodpressure_model that you fit in the previous exercise.

You will also compare the predictions to outcomes graphically. ggplot2 is already loaded in your workspace. Recall the plot command takes the form:

      ggplot(dframe, aes(x = pred, y = outcome)) + 
           geom_point() + 
           geom_abline(color = "blue")

```{r}
# bloodpressure is in your workspace
summary(bloodpressure)

# bloodpressure_model is in your workspace
bloodpressure_model

# predict blood pressure using bloodpressure_model :prediction
bloodpressure$prediction <- predict(bloodpressure_model, bloodpressure)

# plot the results
ggplot(bloodpressure, aes(x = prediction, y = blood_pressure)) +
     geom_point() + 
     geom_abline(color = "blue")
```

-----------------------------